Orienting the causal relationship between imprecisely measured traits using genetic instruments

### 2017-03-21 docs/manuscript.Rmd
updated for submission

>>>>

### Author summary

Understanding the causal relationships between pairs of traits is crucial for unravelling the causes of disease. To this end, results from genome-wide association studies are valuable because if a trait is known to be influenced by a genetic variant then this knowledge can be used to test the trait's causal influences on other traits and diseases. Here we demonstrate that when there is uncertainty about the nature of the genetic association several existing methods that are used for causal inference can be liable to give the wrong direction of causality. We introduce a new method that is more flexible and robust to problems such as measurement error, and apply it to evaluate the causal relationships between DNA methylation levels and gene expression. Our results show that DNA methylation is more likely to be the causal factor, but that systematic differences in measurement error between the platforms could influence such conclusions.

### 2017-03-21 docs/manuscript.tex
updated for submission

\setlength{\droptitle}{-2em}
  \title{Orienting the causal relationship between imprecisely measured traits
using genetic instruments}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{}
  \preauthor{}\postauthor{}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{15 March 2017}

@@

>>>>

\setlength{\droptitle}{-2em}
  \title{Orienting the causal relationship between imprecisely measured traits
using genetic instruments}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{}
  \preauthor{}\postauthor{}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{20 March 2017}

@@

\subsubsection{Author summary}\label{author-summary}

Understanding the causal relationships between pairs of traits is
crucial for unravelling the causes of disease. To this end, results from
genome-wide association studies are valuable because if a trait is known
to be influenced by a genetic variant then this knowledge can be used to
test the trait's causal influences on other traits and diseases. Here we
demonstrate that when there is uncertainty about the nature of the
genetic association several existing methods that are used for causal
inference can be liable to give the wrong direction of causality. We
introduce a new method that is more flexible and robust to problems such
as measurement error, and apply it to evaluate the causal relationships
between DNA methylation levels and gene expression. Our results show
that DNA methylation is more likely to be the causal factor, but that
systematic differences in measurement error between the platforms could
influence such conclusions.
